Community biomass and species richness temporal trends reveal contrasting assembly and disassembly processes in natural food-webs

### 2022-10-24 paper/manuscript.Rmd
Switched all figures and analysis to targets

```{r}
loadd(rsq_table_bm_rich_mono_stable_trends)
loadd(reg_table_bm_rich_mono_stable_trends)
load(here::here("data", "semeff.rda"))
load(here::here("data", "sp_semeff.rda"))
```

```{r}
net_var_bm <- c("log_rich_std", "ct_ff", "w_trph_lvl_avg", "piel_bm",
      "piel_nind")
net_var_rich <- c("log_bm_std", "ct_ff", "w_trph_lvl_avg", "piel_bm",
      "piel_nind")
```

```{r}
loadd(slope_com_var_no_covar, st_mono_trends_stable_rich_bm)
loadd(range_data)
```
```{r}
nb_st <- unique(slope_com_var_no_covar$station) %>% length
rm(slope_com_var_no_covar)
nb_st_rich_bm <- st_mono_trends_stable_rich_bm %>% length
```

@@

# Methods

## Data

## Data selection

To characterize the links among the temporal trends of network and community
structure and the temporal trends of community biomass and species richness, we
characterized the temporal trajectories of food-web structure at each site. We
modeled a linear relationship between food-web structure and
time with a second order polynome, which allows to assess both the direction and the monotony of temporal
trends [@rigal_method_2020]. We selected the sites that displayed monotonic
trends for these variables, i.e. the ones for which the second order polynomial
term was non significant according to t-tests, to avoid the inclusion of many
types of biomass and species richness trends (i.e. accelerating, decelerating,
etc...), . Over the `r nb_st` monitored sites, the selection procedure resulted in 
`r nb_st_rich_bm` sites for community biomass and species richness analysis.
However, the results were robust to the inclusion of non monotonic trends (Fig.
SXX).

There was a clear relationship between the size of the communities and their
temporal dynamics. Communities that displayed low community biomass at the
beginning of their monitoring tend to display a biomass increase through time
and reversely, we observed the same pattern for species richness. We
log-transformed (natural logarithm) community biomass and species richness to model their trends,
which decreased the former effect (Fig. SXX). Temporal
trends of community biomass and species richness were thus expressed in rate of
change by year.

In order to control for the uncertainties of the estimated temporal trends, each
temporal trend was weighted by the inverse of the sum of the standard errors
of the estimated trends, i.e. estimated slopes for the response and the
explicative variables ($w_j = \dfrac{1}{\sum{e_j}}$, $w_j$ being the
weight, and $s_{e_j}$ the standard error on the
temporal trends of the variables).

The coefficients of the Structural Equation Models were computed with
PiecewiseSEM R package [@lefcheck_piecewisesem_2016]. The coefficients were
standardizated and the confidence intervals were computed with semEff R package
[@murphy_semEff_2021]. Confidence intervals were obtained by bootstrapping the
linear models 1000 times and computing adjusted bootstrap percentile interval
(BCa), which are adjusted for asymetric distribution of resampling.

```{r}

fig_mat_cap <- paste0("Using  sites of stream fishes in French streams that
  have been monitored at least 10 years, 1) we infered the food-web at each site
and each sampling event based on ontogenic food diet information for each
species and predator-prey body size ratio. 2) We computed the temporal trends
and the median of community biomass, species richness, connectance, average
trophic level and evenness of biomass at each site. 3) We then assessed the
relationships among a) temporal trends of food-web structure. We also assessed
the spatial relationships among median food-web structure while controlling for
the environmental gradient with two axis of a rotated PCA. The first axis of the
PCA is related to the stream gradient from upstream to downstream while the
second is related to enrichment (BOD) and site altitude.")

```

```{r matmet, fig.cap=fig_mat_cap, out.width="80%"}
knitr::include_graphics(get_mypath("figures", "mat_met_inkscape.pdf"))
```

@@

```{r}
loadd(rsq_sp_mod, summary_table, sp_models, model_summary)

loadd(rsq_table_bm_rich_mono_stable_trends, sp_rsq_table_bm_rich_mono_stable_trends)

loadd(rigal_classification)
```

```{r}
log_trends_percent <- rigal_classification %>%
  unnest(classif) %>%
  filter(variable %in% c("log_bm_std", "log_rich_std")) %>%
  select(variable, station, linear_slope, linear_slope_strd_error) %>%
  mutate(
    slp_percent = round((exp(linear_slope) - 1) * 100),
    std_err_percent = round((exp(linear_slope_strd_error) - 1) * 100)
  )
resume_log_trends <- log_trends_percent %>%
  group_by(variable) %>%
  summarise(
    mean = mean(slp_percent, na.rm = TRUE),
    sd = sd(slp_percent, na.rm = TRUE))
```

```{r}
class_trends_bm_rich <- rigal_classification %>%
  filter(variable %in% c("log_bm_std", "log_rich_std")) %>%
  unnest(classif) %>%
  select(variable, station, direction, acceleration, shape_class)

res_trends <- class_trends_bm_rich %>%
  filter(acceleration == "constant") %>%
  group_by(variable, direction) %>%
  summarise(n = n()) %>%
  group_by(variable) %>%
  mutate(total = sum(n), percent = round(n / sum(n) * 100))

x_var <- c("log_bm_std", "log_rich_std")
stable_st_var <- map(x_var,
  ~class_trends_bm_rich[
    class_trends_bm_rich$variable == .x &
    class_trends_bm_rich$direction == "stable",
      ]$station
)
names(stable_st_var) <- x_var

#sum(stable_st_var[["log_bm_std"]] %in% stable_st_var[["log_rich_std"]] == TRUE) / length(stable_st_var[["log_bm_std"]])

#sum(stable_st_var[["log_rich_std"]] %in% stable_st_var[["log_bm_std"]] == TRUE) / length(stable_st_var[["log_rich_std"]])
```

```{r}
loadd(std_table_bm_rich_mono_stable_trends)
std <-
  std_table_bm_rich_mono_stable_trends %>%
  mutate(
    type = recode(type, spatial = "sp", temporal = "tps"),
    response = recode(response, log_bm_std = "bm", log_rich_std = "rich", ct_ff
      = "ct", piel_bm = "p_bm", piel_nind = "p_n", w_trph_lvl_avg = "tlvl"),
    term = recode(term, log_bm_std = "bm", log_rich_std = "rich")
    ) %>%
  rename(t = type, rep = response, sest = std_estimate) %>%
  mutate_at(vars(sest, stdse), ~round(.x, 2))
tps_bm <- std %>%
  filter(t == "tps", term == "bm")

tps_bm_sest <- tps_bm$sest
names(tps_bm_sest) <- tps_bm$rep
```

```{r}
tps_r <- std %>%
  filter(t == "tps", term == "rich")

tps_r_sest <- tps_r$sest
names(tps_r_sest) <- tps_r$rep
```

```{r}
sp_b <- std %>%
  filter(t == "sp", term == "bm")
sp_b_sest <- sp_b$sest
names(sp_b_sest) <- sp_b$rep

sp_r <- std %>%
  filter(t == "sp", term == "rich")
sp_r_sest <- sp_r$sest
names(sp_r_sest) <- sp_r$rep
```

```{r}
rsq_ct <- rsq_table_bm_rich_mono_stable_trends %>%
  filter(Response == "ct_ff") %>%
  .$R.squared %>%
  round(., 2)

rsq_bm <- rsq_table_bm_rich_mono_stable_trends %>%
  filter(Response == "log_bm_std") %>%
  .$R.squared %>%
  round(., 2)

rsq_tlvl_sp <- sp_rsq_table_bm_rich_mono_stable_trends %>%
  filter(Response == "w_trph_lvl_avg") %>%
  .$Conditional %>%
  round(., 2)

rsq_ct_sp <- sp_rsq_table_bm_rich_mono_stable_trends %>%
  filter(Response == "ct_ff") %>%
  .$Conditional %>%
  round(., 2)
```
Over the sampling sites in the dataset, we found that
`r res_trends[res_trends$variable == "log_bm_std", ]$total[1]` stations displayed monotonic trends for biomass. Among them, 
`r res_trends[res_trends$variable == "log_bm_std" & res_trends$direction == "stable", ]$percent`%
were classified as stable,
`r res_trends[res_trends$variable == "log_bm_std" & res_trends$direction == "decrease", ]$percent`%
as decreasing,
`r res_trends[res_trends$variable == "log_bm_std" & res_trends$direction == "increase", ]$percent`%
as increasing. Regarding the species richness, we found that
`r res_trends[res_trends$variable == "log_rich_std", ]$total[1]` stations displayed monotonic trends.
`r res_trends[res_trends$variable == "log_rich_std" & res_trends$direction == "stable", ]$percent`%
were classified as stable,  
`r res_trends[res_trends$variable == "log_rich_std" & res_trends$direction == "decrease", ]$percent`%
as decreasing, and 
`r res_trends[res_trends$variable == "log_rich_std" & res_trends$direction == "increase", ]$percent`% as increasing.

Overall, we found relatively low in the explicative power of statistical
models linking temporal dynamics of food-web structure and the one of community
biomass and species richness spanning from $R_{adj}^2= `r rsq_ct`$ for
connectance to $R_{adj}^2= `r rsq_bm`$ for community biomass (Table SXX).
The models over the spatial gradient were more performant, with
$R_{cond}^2$ varing from $`r rsq_tlvl_sp`$ for the average trophic level to
$`r rsq_ct_sp`$ for connectance (Table SXX), $R_{cond}^2$ being the
$R^2$conditional to the random effect.

---
#Overall, we found no significant effect of the temporal gain or losses of
#communinity biomass biomass and species richness on the relationship between the
#temporal trends of community structure and that of biomass and species richness
#(Table \@ref(tab:aov)).
---

```{r}
loadd(sp_fig1_2_bm_rich_mono_stable_trends,
  pred_bm_rich_mono_stable_trends)

fig1_2 <- get_plot_rich_bm(
  predict_plot = pred_bm_rich_mono_stable_trends,
  get_list = TRUE,
  rm_legend = FALSE,
  bm_x = "log_bm_std",
  bm_y = c("ct_ff", "w_trph_lvl_avg"), #, "piel_nind"
  rich_y = c("ct_ff", "w_trph_lvl_avg", "log_bm_std"), #, "piel_nind"
  rich_x = "log_rich_std"
  )
p_bm <- fig1_2$bm

p_sp_bm <- sp_fig1_2_bm_rich_mono_stable_trends$bm[c("log_rich_std", "ct_ff", "w_trph_lvl_avg")]
```

```{r}
fig_bef_cap <- paste0(
  "Relationship between the temporal trends of community biomass 
  and the one of species richness. Red line correspond to the predictions of the
  linear models. The lines surrounding points display the standard error
  associated to the temporal trends. The histograms display the distribution of
  response and predictive variables."
)
```

```{r}
library(ggExtra)
coeff_bef <- c(
  rsq = rsq_table_bm_rich_mono_stable_trends %>%
    filter(Response == "log_bm_std") %$%
    unlist(.[, c("R.squared")]),
  reg_table_bm_rich_mono_stable_trends %>%
    filter(response == "log_bm_std", term == "log_rich_std") %$%
    unlist(.[, c("estimate", "std.error", "std_estimate", "stdse")])
  ) %>%
round(., 2)

beta_rsq <- function(
  rsq = NULL,
  beta = NULL,
  se = NULL) {

  if (!is.null(rsq)) {

    tmp <- "paste(
    beta, \" (s.e.) \", \" = \",
    mybeta, \" (\", myse, \"), \",
    italic(R)^2, \" = \", myrsq
    )"
str_replace_all(tmp,
  c(
    "myrsq" =  as.character(rsq),
    "mybeta" = as.character(beta),
    "myse" = as.character(se)
  )
)

  } else {

    tmp <- "paste(
    beta, \" (s.e) \", \" = \",
    mybeta, \" (\", myse, \") \" )"
str_replace_all(tmp,
  c(
    "mybeta" = as.character(beta),
    "myse" = as.character(se)
  )
)

  }

}

scale_xy_log_trends <- function(breaks = NULL, axis = "y", sep ="\n") {
  log_to_percent <- round((exp(breaks) - 1) * 100)
  labels <- paste0(breaks, sep, log_to_percent, " %")

  if(axis == "y") {
    fun <- scale_y_continuous
  } else {
    fun <- scale_x_continuous
  }

  fun(breaks = breaks, labels = labels)

}
```

We found a strong positive relationship between the temporal trends of community
biomass and the ones of species richness ($r_\delta = `r tps_r_sest["p_bm"]`$,
Fig. \@ref(fig:bmr) B), meaning that both temporal trends
had the same direction in average. At the sites 
where species richness decreased over time, community biomass also decreased
over time. On the other hand, at the sites where species richness increased over
time, community biomass also increased over time. Along the spatial gradient, we also
found a positive relationship between median community biomass and median
species richness, but half the strength as the temporal one
($r_\delta = `r sp_b_sest["rich"]`$).

The temporal trends of connectance were positively linked with the ones of
community biomass ($r_\delta = `r tps_bm_sest["ct"]`$), also meaning that both
temporal trends had the same direction in average. That is, the sites where
food-web connectance decreased over time also had community biomass that
decreased over time and reversely. It suggests that communities that decreased
in biomass over time loose generalist predators over time, as connectance
measures the degree of generalism, and reversely that communities
gaining biomass over time also gain generalist predators. Spatially, we also found
a positive relationship between median connectance and median species
richness ($r_\delta = `r sp_b_sest["ct"]`$). We found a negative 
relationship between the temporal trends of connectance and the
ones of species richness ($r_\delta = `r tps_r_sest["ct"]`$, Fig. \@ref(fig:bmr)D)
while we found negative relationship three times stronger between median connectance and
median species richness along the spatial gradient 
($r_\delta = `r sp_r_sest["ct"]`$, Fig. \@ref(fig:std)).

The temporal trends of average trophic level had a strong positive
relationship with the ones of community biomass 
($r_\delta = `r tps_bm_sest["tlvl"]`$, Fig. \@ref(fig:bmr)B), meaning
that sites where community biomass decreased over time, average trophic level
decreased over time and *vice versa*. It means that in communities loosing biomass
over time, the average biomass of the food-web moved down to lower trophic level
and *vice versa*. We also found a positive relationship between
median average trophic level and median community biomass in space, but half the
strength as the temporal one ($r_\delta = `r sp_b_sest["tlvl"]`$, Fig.
\@ref(fig:std)). The temporal trends of average trophic level was unrelated to
the ones of species richness. It suggests that in communities either loosing or
gaining species richness over time, the location of average biomass along
trophic level did not significantly change according to species richnnes.
However, we found a strong negative relationship in space 
($r_\delta = `r sp_r_sest["tlvl"]`$, Fig. \@ref(fig:std)).

```{r}
sp_semeff_tab <- from_semEff_to_table(sp_semeff)
semeff_tab <- from_semEff_to_table(semeff)
semeff_tot <- rbind(
  tibble(semeff_tab, model = "temporal"),
  tibble(sp_semeff_tab, model = "spatial")
  ) %>%
filter(
  !predictor %in% c("log_RC1", "log_RC2"),
  effect_type != "mediators"
)
```

```{r}
semeff_tab_coeff <- semeff_tab %>%
  mutate(across(where(is.numeric), ~round(., 2))) %>%
  mutate(eci = paste0(effect, " [", lower_ci, ";", upper_ci, "]"))

sp_semeff_tab_coeff <- sp_semeff_tab %>%
  mutate(across(where(is.numeric), ~round(., 2))) %>%
  mutate(eci = paste0(effect, " [", lower_ci, ";", upper_ci, "]")) %>%
  filter(! predictor %in% c("log_RC1", "log_RC2"))

ctsi <- semeff_tab_coeff %>%
  filter(response == "ct_ff", effect_type == "indirect")
ctstr <- semeff_tab_coeff %>%
  filter(response == "ct_ff", effect_type == "total",
    predictor == "log_rich_std")

ctspi <- sp_semeff_tab_coeff %>%
  filter(response == "ct_ff", effect_type == "indirect")
ctsptr <- sp_semeff_tab_coeff %>%
  filter(response == "ct_ff", effect_type == "total",
    predictor == "log_rich_std")

wtsi <- semeff_tab_coeff %>%
  filter(response == "w_trph_lvl_avg", effect_type == "indirect")
wtstr <- semeff_tab_coeff %>%
  filter(response == "w_trph_lvl_avg", effect_type == "total",
    predictor == "log_rich_std")

wtspi <- sp_semeff_tab_coeff %>%
  filter(response == "w_trph_lvl_avg", effect_type == "indirect")
wtsptr <- sp_semeff_tab_coeff %>%
  filter(response == "w_trph_lvl_avg", effect_type == "total",
    predictor == "log_rich_std")
```

The structural model on the temporal trends of food-web structure reveals that
the species richness has a direct negative effect on connectance but also a
negative one, twice lower than the former, through community biomass
($r_\delta$ = `r ctsi[["eci"]]`). It resulted a relatively weak total negative
effect of species richness on connectance (`r ctstr[["eci"]]`). Spatially,
we also found a positive indirect effect of richness on connectance through
community biomass but a weak one (`r ctspi[["eci"]]`), resulting in a strong negative total
effect of richness on connectance (`r ctsptr[["eci"]]`).

The relationships among temporal trends showed a indirect positive effect of
species richness on average trophic level (`r wtsi[["eci"]]`), resulting in a total positive
effect of species richness on average trophic level (`r wtstr[["eci"]]`). Over the spatial
gradient, we found the same positive indirect effect of richness through
community biomass (`r wtspi[["eci"]]`). But as the direct effect of richness on connectance
was strongly negative, it resulted in a total strong negative effect of species
richness on connectance over the spatial gradient (`r wtsptr[["eci"]]`).

```{r}
fig_bm_cap <- paste0(
  "Relationship between community structure and community biomass. Left panel
  displays the relationships between the temporal trends of community structure
  and the one of community biomass. Lines correspond to the predictions of the
  linear models while the points are colored according the median community
  biomass of the three first years of the given temporal series. The lines
  surrounding each point represent the standard error associated to the temporal
  trends. Right panel displays the
  spatial relationship between median community structure over time and the
  median community biomass over time."
)

fig_rich_cap <- paste0(
  "Relationship between community structure and species richness. Left panel
  displays the relationships between the temporal trends of community structure
  and the one of species richness. Lines correspond to the predictions of the
  linear models while the points are colored according the median species
  richness of the three first years of the given temporal series. The lines
  surrounding each point represent the standard error associated to the temporal
  trends. Right panel displays the spatial relationship between median community
  structure over time and the median species richness over time."

)
```

```{r}
fig_bm_rich_cap <- paste0(
  "Relationships between the temporal trends of food-web structure
  and the ones of community biomass (left panel, A, C, E) species richness (right panel, B, D, F).
  The lines surrounding each point display the standard errors associated to
  the estimated temporal trends. Red and black lines represent respectively
  significant and non-significant slopes, both lines displays prediction from
  the Ordinary Least Squares model. The histograms display the distribution of response and predictive
  variables. Labels display the unstandardised slope coefficients and their
  standard errors (see Fig. \\ref{fig:std} for the standardized coefficients).
  $R^2$ are displayed only on the right panel, as it is the same model for both
  panels.")

```

```{r bef, fig.cap=fig_bef_cap, out.width="50%"}
lab_log_rich <- expression(atop("Log species richness ("*m^{-2}*")","trend slope (log / % per year)"))
lab_log_bm <- expression(atop("Log community biomass ("*g.m^{-2}*")", "trend slope (log / % per year)"))

p_bef <- fig1_2$rich$log_bm_std +
  labs(
    y = lab_log_bm,
    x = lab_log_rich
  ) +
scale_xy_log_trends(breaks = c(-.10, -.050, 0, .050, .10), axis = "x") +
scale_xy_log_trends(breaks = c(-.4, -.2, 0, .4, .2), axis = "y", sep = " / ") +
    theme(
      axis.text = element_text(size = 8),
      axis.title = element_text(size = 10)
      )

```

```{r}
fig2 <- fig1_2
fig2$rich$log_bm_std <- NULL
fig2$rich <- map(fig2$rich,
  ~.x +
    theme(
      axis.text = element_text(size = 8),
      axis.title = element_text(size = 10)
      )
  )
fig2$bm <- map2(fig2$bm, names(fig2$rich),
  ~ .x + theme(
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 10)
    ) +
  labs(
    y = paste0(var_replacement()[.y],
      "\ntrends (per year)"),
    x = lab_log_bm
  ) +
  scale_xy_log_trends(breaks = c(-.40, -.20, 0, .20, .40), axis = "x")
)
# Delete redondant legend
fig2$bm[[1]] <- fig2$bm[[1]] + theme(
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title.x = element_blank()
  )

fig2$rich <- map(fig2$rich,
  ~ .x + theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank()
  ) +
  labs(x = lab_log_rich) +
  scale_xy_log_trends(breaks = c(-.10, -.050, 0, .050, .10), axis = "x")
  )
fig2$rich[[1]] <- fig2$rich[[1]] + theme(
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title.x = element_blank()
  )

```

```{r, eval = FALSE}
# Add beta and Rsq
for (i in seq_along(fig2$rich)) {

  resp_var <- names(fig2$rich)[i]

  coeff_rich <- c(
    rsq = rsq_table_bm_rich_mono_stable_trends %>%
      filter(Response == resp_var) %$%
      unlist(.[, c("R.squared")]),
    reg_table_bm_rich_mono_stable_trends %>%
      filter(response == resp_var, term == "log_rich_std") %$%
      unlist(.[, c("estimate", "std.error", "std_estimate", "stdse")])
    ) %>%
  round(., 2)

  coeff_bm <- reg_table_bm_rich_mono_stable_trends %>%
    filter(response == resp_var, term == "log_bm_std") %$%
    unlist(.[, c("estimate", "std.error", "std_estimate", "stdse")]) %>%
    round(., 2)

  # Add coeff:
  fig2$rich[[resp_var]] <- fig2$rich[[resp_var]] +
  annotate("text", x = -.10, y = Inf, vjust = 1, hjust = 0,
    size = 2,
    parse = T,
    label = beta_rsq(
      rsq = coeff_rich["rsq"],
      beta = coeff_rich["estimate"],
      se = coeff_rich["std.error"]
    )
  )
  fig2$bm[[resp_var]] <- fig2$bm[[resp_var]] +
  annotate("text", x = -.4, y = Inf, vjust = 1, hjust = 0,
    size = 2,
    parse = T,
    label = beta_rsq(
      rsq = NULL,
      beta = coeff_bm["estimate"],
      se = coeff_bm["std.error"]
    )
  )
}
```

```{r}
# Add histograms
p_bef2 <- p_bef +
  theme(
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title.x = element_blank()
  ) 
p_bef_hist <- ggMarginal(p_bef2, type = "histogram")
#fig2$bm[[1]] <- ggMarginal(fig2$bm[[1]], type = "histogram", margins = "x")
fig2$rich[[1]] <- ggMarginal(fig2$rich[[1]], type = "histogram", margins = "y")
fig2$rich[[2]] <- ggMarginal(fig2$rich[[2]], type = "histogram", margins = "y")
```

```{r}
sem_cap <- "Structural equation models on the temporal trends of the food-web
structure. The path display the standardized slopes and their associated 95%
  confidence intervals."
```

```{r sem, fig.cap = sem_cap}
ti <- magick::image_read(here::here("figures", "sem_tps_avg_weight_ok.svg"),  density = 300)
p <- ggdraw() +
  draw_image(ti,
    x = 0, y = 0, hjust = 0, vjust = 0, width = .9, height = 1)
```

```{r bmr, out.width="100%", fig.height=7, fig.cap=fig_bm_rich_cap}
#p <- draw_image(here::here("figures", "sem_tps_avg_weight_ok.pdf"),
#  x = 1, y = 1, hjust = 1, vjust = 1, width = 0.13, height = 0.2)
biplot_model <- plot_grid(
  plot_grid(NULL, p_bef_hist,
    labels = c("", "A"), ncol = 2,
    rel_widths = c(.54, 1)
    ),
  plot_grid(
  plot_grid(plotlist = fig2$bm,
    ncol = 1, labels = c("B", "D"),
    hjust = -5.5, vjust = 1.75,
    rel_heights = c(.85, 1)
    ),
  plot_grid(plotlist = fig2$rich,
    ncol = 1, labels = c("C", "E"),
    hjust = -2, vjust = 1.75,
    rel_heights = c(.85, 1)
    ),
  ncol = 2,
  rel_widths = c(.9, 1)
  ),
  nrow = 2,
  rel_heights = c(.55, 1)
)
sem_p <- plot_grid(p, biplot_model, ncol = 2, rel_widths = c(.4, .6))

save_plot(
  filename = here::here("paper", "figures", "biplot_tps.png"),
  biplot_model, nrow = 3, ncol = 2, base_asp = 1)
save_plot(
  filename = here::here("paper", "figures", "sem_p.png"),
  sem_p, nrow = 3, ncol =3)
sem_p
```

```{r}
std_cap <- paste0(
  "Assessment of the space-for-time hypothesis. Standardized coefficients
  ($r_\\delta$) derived from the spatial and temporal Structural Equation
  Models. The left, middle and right panels display respectively the direct,
  indirect and total effects of community biomass and species richness on the
  Average trophic level (top panel), the connectance (middle panel) and the
  community biomass (bottom panel) The lines display the confidence intervals at
  95%.")

```

```{r std, fig.cap=std_cap, fig.width=7, fig.asp = .7}
p_sem_eff <- semeff_tot %>%
  mutate(
    response = var_replacement()[response],
    predictor = var_replacement()[predictor],
    effect_type = str_to_sentence(effect_type),
    model = str_to_sentence(model)
    ) %>%
  ggplot(aes(y = predictor, x = effect, colour = effect_type, shape = model)) +
  geom_point(
    position = position_dodge(width = 0.4), width = .4) +
  geom_vline(xintercept = 0) +
  geom_linerange(
    aes(xmin = lower_ci, xmax = upper_ci),
    position = position_dodge(width = 0.4)
    ) +
  facet_grid(rows = vars(response)) +
  labs(
    y = "Predictor",
    x = expression(paste("Standardized slope coefficients ", r[delta])),
    color = "Effect type", shape = "Model") +
  theme(legend.position = "bottom")
p_sem_eff
```

@@

>>>>

```{r}
tar_load(full_data2)
nb_st <- unique(full_data2$station) %>% length
rm(full_data2)
```

# Abstract (250 words max)

# Signifiance (120 words)

@@

@@

```{r}
tar_load(glob_tps_trends_decade_no_drivers)
gbm <- get_effect_ci(resp = "log_bm_std", ci_lvl = "level:0.80", p = TRUE)
gr <- get_effect_ci(resp = "log_rich_std", ci_lvl = "level:0.80", p = TRUE)
gt <- get_effect_ci(resp = "w_trph_lvl_avg", ci_lvl = "level:0.80", p = FALSE)
gct <- get_effect_ci(resp = "ct_ff", ci_lvl = "level:0.80", p = FALSE)
```

```{r}
tar_load(r2_inla_tab)
r2_inla_report <- r2_inla %>%
  select(response, r2_mvp_marg95hpdci, r2_mvp_cond95hpdci)
r2ic <- map(
  setNames(r2_inla_report$response, r2_inla_report$response),
  ~r2_inla_report[r2_inla_report$response == .x, ]$r2_mvp_cond95hpdci
)

tar_load(r2_sem)
```
Using a hierarchical model to estimate conjointly the temporal trends of  
foodweb strucure at each site, we found weak evidence of a global decrease in
community biomass (CI80%:`r gbm` by decade; median [Credible Interval]) but of a
global increase in species richness (CI80%: `r gr` by decade) across the `r nb_st` sites. 
We did not find evidence of a global trends in connectance and average trophic
level (CI80%: `r gct` and `r gt` respectively). The hierarchical model captured
well the variance associated to the spatial structure of the data with
conditional $R^2$ going from  `r r2ic$log_bm_std` (mean [CI95%]) for community
biomass to `r r2ic$log_rich_std` for species richness (Table SXX).

```{r}
sem_cap <- "(a) Structural equation models linking the temporal trends of the food-web
structure. The path widths are proportional to the standardized slope coefficients. The standardized slope coefficients and their associated 95%
  confidence intervals are displayed along the path. (b to f) Direct bivariate
relationships derived from the structural equation model. The distribution of the
site temporal trends are displayed in Fig. SXX."
```

```{r sem, fig.cap = sem_cap}
knitr::include_graphics(here::here("figures", "fig2.png"))
```

```{r}
tar_load(semeff_tot)
semeff_tot <- semeff_tot %>%
  mutate(across(c(response, predictor), ~str_remove(.x, "mean_")))

rdtps <- map(
  setNames(
    c("ct_ff", "w_trph_lvl_avg", "log_bm_std"),
    c("ct", "tlvl", "bm")
    ),
  ~ list(
    bm = get_semeff_ci(resp = .x, term = "log_bm_std"),
    r = get_semeff_ci(resp = .x, term = "log_rich_std")
    )
)
rdsp <- map(
  setNames(
    c("ct_ff", "w_trph_lvl_avg", "log_bm_std"),
    c("ct", "tlvl", "bm")
    ),
  ~ list(
    bm = get_semeff_ci(resp = .x, term = "log_bm_std", model = "spatial"),
    r = get_semeff_ci(resp = .x, term = "log_rich_std", model = "spatial")
    )
)
```

We found a strong positive relationship between the temporal trends of community
biomass and the ones of species richness ($r_\delta = `r rdtps$bm$r`$,
Fig. \@ref(fig:sem) b), meaning that both temporal trends
had the same direction in average. At the sites 
where species richness decreased over time, community biomass also decreased
over time. On the other hand, at the sites where species richness increased over
time, community biomass also increased over time. Along the spatial gradient, we also
found a positive relationship between median community biomass and median
species richness, but half the strength as the temporal one
($r_\delta = `r rdsp$bm$r`$).

The temporal trends of connectance were positively linked with the ones of
community biomass ($r_\delta = `r rdtps$ct$bm`$), also meaning that both
temporal trends had the same direction in average. That is, the sites where
food-web connectance decreased over time also had community biomass that
decreased over time and reversely. It suggests that communities that decreased
in biomass over time loose generalist predators over time, as connectance
measures the degree of generalism, and reversely that communities
gaining biomass over time also gain generalist predators. Spatially, we also found
a positive relationship between median connectance and median species
richness ($r_\delta = `r rdsp$ct$bm`$). We found a negative 
relationship between the temporal trends of connectance and the
ones of species richness ($r_\delta = `r rdtps$ct$r`$, Fig. \@ref(fig:sem)d)
while we found negative relationship three times stronger between median connectance and
median species richness along the spatial gradient 
($r_\delta = `r rdsp$ct$r`$, Fig. \@ref(fig:std)).

The temporal trends of average trophic level had a strong positive
relationship with the ones of community biomass 
($r_\delta = `r rdtps$tlvl$bm`$, Fig. \@ref(fig:sem)B), meaning
that sites where community biomass decreased over time, average trophic level
decreased over time and *vice versa*. It means that in communities loosing biomass
over time, the average biomass of the food-web moved down to lower trophic level
and *vice versa*. We also found a positive relationship between
median average trophic level and median community biomass in space, but half the
strength as the temporal one ($r_\delta = `r rdsp$tlvl$bm`$, Fig.
\@ref(fig:std)). The temporal trends of average trophic level was unrelated to
the ones of species richness. It suggests that in communities either loosing or
gaining species richness over time, the location of average biomass along
trophic level did not significantly change according to species richnnes.
However, we found a strong negative relationship in space 
($r_\delta = `r rdsp$tlvl$r`$, Fig. \@ref(fig:std)).

```{r}
ridtps <- map(
  setNames(
    c("ct_ff", "w_trph_lvl_avg", "log_bm_std"),
    c("ct", "tlvl", "bm")
    ),
  ~ list(
    r = get_semeff_ci(resp = .x, term = "log_rich_std", type = "indirect")
    )
)
ridsp <- map(
  setNames(
    c("ct_ff", "w_trph_lvl_avg", "log_bm_std"),
    c("ct", "tlvl", "bm")
    ),
  ~ list(
    r = get_semeff_ci(resp = .x, term = "log_rich_std",
      model = "spatial", type = "indirect")
    )
)
rttps <- map(
  setNames(
    c("ct_ff", "w_trph_lvl_avg", "log_bm_std"),
    c("ct", "tlvl", "bm")
    ),
  ~ list(
    bm = get_semeff_ci(resp = .x, term = "log_bm_std", type = "total"),
    r = get_semeff_ci(resp = .x, term = "log_rich_std", type = "total")
    )
)
rtsp <- map(
  setNames(
    c("ct_ff", "w_trph_lvl_avg", "log_bm_std"),
    c("ct", "tlvl", "bm")
    ),
  ~ list(
    bm = get_semeff_ci(resp = .x, term = "log_bm_std",
      model = "spatial", type = "total"),
    r = get_semeff_ci(resp = .x, term = "log_rich_std",
      model = "spatial", type = "total")
    )
)
```

The structural model on the temporal trends of food-web structure reveals that
the species richness has a direct negative effect on connectance but also a
positive one through community biomass
($r_\delta$ = `r ridtps$ct$r`). It resulted a relatively weak total negative
effect of species richness on connectance (`r rttps$ct$r`). Spatially,
we also found a positive indirect effect of richness on connectance through
community biomass but a weak one (`r ridsp$ct$r`), resulting in a strong negative total
effect of richness on connectance (`r rtsp$ct$r`).

The relationships among temporal trends showed a indirect positive effect of
species richness on average trophic level (`r ridtps$tlvl$r`), resulting in a total positive
effect of species richness on average trophic level (`r rttps$tlvl$r`). Over the spatial
gradient, we found the same positive indirect effect of richness through
community biomass (`r ridsp$tlvl$r`). WHATTTT? But as the direct effect of
richness on average trophic level was strongly negative, it resulted in a total
strong negative effect of species richness on average trophic level over the
spatial gradient (`r rtsp$tlvl$r`).

```{r}
std_cap <- paste0(
  "Assessment of the space-for-time hypothesis. Standardized coefficients
  ($r_\\delta$) derived from the spatial and temporal Structural Equation
  Models. The left, middle and right panels display respectively the direct,
  indirect and total effects of community biomass and species richness on the
  Average trophic level (top panel), the connectance (middle panel) and the
  community biomass (bottom panel) The lines display the confidence intervals at
  95%.")
```

```{r std, fig.cap=std_cap, fig.width=7, fig.asp = .7}
knitr::include_graphics(here::here("figures", "p_sem_tps_sp.pdf"))
```

@@

# Material & Methods

We modeled the temporal trends of food-web structure at each site ($Y_i$) using 
linear hierarchical model. Time was considered as the number of year since the
first sampling ($t = 0$) at each site. We accounted for the spatial structure of
the data by adding random intercepts ($a$) and slopes ($b$) for the
basin ($a_n$, $b_n$) and site ($a_{i|n}$, $b_{i|n}$) identity, site effects
being nested in the basin one.

$$
\begin{aligned}
    Y_{i|n, t} &= \alpha + \betaTime_t + \epsilon_{i|n, t}
  (\#eq:gen)
\end{aligned}
$$

- $\alpha = \alpha_0 + a_n + a_{i|n}$
- $\beta = \mu + b_n + b_{i|n}$
- $a_n, a_{i|n}, b_n, b_{i|n}, \epsilon_{i|n, t} \sim \mathcal{N}(0,\,\sigma^{2})$
- $n$: basin, $i$: site $i$, $t$: time $t$

The model parameters were estimated using INLA (Integrated Nested Laplacian
Approximation) bayesian inference with default uninformative priors
[@rue_bayesian_2017]. The prior distributions for the fixed intercept and slope were
modelled as a flat normal distribution with a mean of zero ($\mathcal{N}(0,\, 1000)$).
Prior distributions for the gaussian error $\epsilon_{i|n, t}$ and the random
effects were modelled as a log gamma distribution with
shape and inverse scale parameters ($\mathcal{G}(s, \, \tau) = \mathcal{G}(1, \,
5.10^-5)$), then their respective standard deviation were obtained by
backtransforming the inverse scale parameter ($\sigma = 1 / \sqrt{\tau}$).
Community biomass and species richness were logged transformed.
